Switching Optimisation in Huffman Code for Power Efficient Data Transmission

### 2015-02-25 Switche Paper 1st approache/elsarticle-template-num.tex
abstract

\title{Reducing Power Consumption in Huffman Coding}

@@

\begin{abstract}
%% Text of abstract
Recently the power consumption identified as one of the most pressing challenges of data transmission. Though a considerable reduction of size of data is obtained by Huffman code and several methods have been proposed to improve the data compression rate. These methods improves only the total number of bits without consideration of transmission cost. In this paper, we propose a new approach to reduce the power consumption, by applying genetic algorithm for minimising the total number of switches, either inter- or intra-switch. 
The approach starts it operation by generating a set of trees considered as initial population for the genetic algorithm, each one of these trees represent a set of codewords (solution), after that the genetic operators such as selection, crossover and mutation will be applied to the initial population in order to evolute the quality of the solutions. The amount of power consumption is proportional to the total number of switches on the whole sequence. 
The performance of the approach is evaluated by applying it to compress some standard biological datasets. The experiments yield that the proposed approach improves the power consumption rate considerably comparing with classic Huffman code.
\end{abstract}

@@

The rest of the paper is organised as follows:
Section 2 presents the background study of the
relationship of switching activity with power consumption, the relation of switching with Huffman code and the issues in biological data transmission. The proposed approach is described in Section 3. Experimental results and
discussion are presented in Section 4 . Finally,
concluding remarks are presented in Section 5. 
\section{Background}
\subsection{Huffman Coding}
\subsection{Power consumption}

>>>>

\title{Switching Optimisation in Huffman Code for Power Efficient   Data Transmission }

@@

\begin{abstract}

Reliability of low powered devices is highly dependent on their power efficiency. In digital communication, a significant amount of power is dissipated in data transmission. Different technologies have been emerged to address the issues of power consumption. In CMOS technology, dynamic power accounts for 70\%-90\% of the total power dissipation and it depends on the representation of the data  and increases linearly with switching activities (transition from logic level High to Low and vice versa). Therefore, an efficient representation of data can minimise power consumption by reducing switching activities.  In this paper, we have extended the Huffman code, a widely used data compression technique, by using genetic algorithm to reduce switching activities in the transmitted message. %In Recently the power consumption identified as one of the most pressing challenges of data transmission. Though a considerable reduction of size of data is obtained by Huffman code and several methods have been proposed to improve the data compression rate. These methods improves only the total number of bits without consideration of transmission cost. In this paper, we propose a new approach to reduce the power consumption, by applying genetic algorithm for minimising the total number of switches, either inter- or intra-switch. 
The main objective of the proposed approach is to minimise the switching activities in the codeword of each symbol as well as the switching activities between the symbols.  The approach starts its operation by generating an initial population, a set of Huffman trees, for all the input symbols. Afterwards, the genetic operators such as selection, crossover and mutation are applied to the initial population to improve the quality of the solutions. %The amount of power consumption is proportional to the total number of switches on the whole sequence. 
The performance of the approach is evaluated by applying it to a set of real biological datasets. The experiments yield that the proposed approach reduces the switching activity by $45.47\%$ in the best case, by $36.33\%$ in the average case and by $16.42\%$ in the worst case. 
\end{abstract}

@@

The rest of the paper is organised as follows:
Section 2 presents the background study of the
relationship of switching activity with power consumption, the relation of switching with Huffman code and the issues in biological data transmission. The proposed approach is described in Section 3. Experimental results and
discussion are presented in Section 4 . Finally,
concluding remarks are presented in Section 5. 
\section{Background}
\subsection{Effect of Switching in Power Consumption}
\subsection{Huffman Code and Its Relationship with Switching Activity}
%\subsubsection{Relationship between switching and Huffman Code}
